Parameterized convection, stratiform clouds and resolution sensitivity in the Community Atmosphere Model

### 2019-08-28 resolution.tex
edits

Establishing a complete understanding of resolution sensitivity in atmospheric general circulation models (AGCMs) is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of scale-aware physics. This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception (see Figure~\ref{fig:cam-history}). As with most versions of CAM, in CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution. Prior studies have established that the drying and reduction in cloudiness is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advect dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,YETAL2014JCLIM,HR2017JCLIM}. This study aims to then understand why and how vertical velocities change with resolution, and why the convectiive precipitation decreases at the expense of stratiform precipitation with increasing resolution.

>>>>

Establishing a complete understanding of resolution sensitivity in atmospheric general circulation models (AGCMs) is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of scale-aware physics. This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception (see Figure~\ref{fig:cam-history}).

As with most versions of CAM, in CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution. Prior studies have established that the drying and reduction in cloudiness is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advect dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,YETAL2014JCLIM,HR2017JCLIM}. This study aims to then understand why and how vertical velocities change with resolution, and why convective precipitation decreases at the expense of stratiform precipitation with increasing resolution.
